A saturation model for impaired learning with enhanced plasticity: Supplementary material

### 2013-10-01 MHCpaper/MHCsupp.tex
typos, ref and name of MHC mutant

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
%\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\ko}{_{\text{MHC}}}
\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

The evolution of this probability is described by a forgetting matrix, $\frg$:
%
\begin{equation}\label{eq:evolve}
  \diff{\pr(t)}{t} = r\pr(t)\frg,
  \qquad
  \frg = f\pot\M\pot+f\dep\M\dep-\I,
\end{equation}
%
where $\I$ is the identity matrix.
Eventually, this will settle into the equilibrium distribution:
%
\begin{equation}\label{eq:eqprob}
  \eq\frg=0.
\end{equation}
%

@@

The MHC mutant has a lower threshold for depression.
We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
We will use a larger value for $q$ for depression in the \KO\ mutant.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
We will use a larger value for $x$ for depression in the \KO\ mutant.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
We will use larger values for depression in the \KO\ mutant.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{tab:params}.

@@

This model is parameterised by the values of $x$ chosen for potentiation and depression.
We will use a larger value for depression in the \KO\ mutant.
The values used are listed in \autoref{tab:params}.

@@

Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
It will change the environment, which will lead to a change in how often potentiation and depression occur.
It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.

@@

We will assume $f\dep\wt=f\dep\ko$.
This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
Therefore the rates of potentiation and depression should not change very much either.
This implies that the mean synaptic weight in equilibrium differs between the mutant and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and mutant for the reasons discussed above.
We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{tab:params}.

We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.

@@

The features of the actual experiments that we'd like to capture are:
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
\end{enumerate}
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.

We will not worry too much about the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
However, both the wild-type and mutant should be affected in the same way, so the difference in the curvature of the mutant and wild-type curves after pre-training would be nice to capture.

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.

@@

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
    \cline{1-7}
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
    Pooled        & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
    Non-uniform MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. }

@@

Now, consider the mutant, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.

However, this is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.

Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the mutant (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
This will last for a shorter time in the mutant, due to the enhanced transitions.
This will not be a feature of any of the other models that we study here.
Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
The difference in these two curvatures is therefore a good feature of this model.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^{\mathrm{b}-/-}$}
\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

The evolution of this probability is described by a continuous time Markov matrix matrix, $\frg$:
%
\begin{equation}\label{eq:evolve}
  \diff{\pr(t)}{t} = r\pr(t)\frg,
  \qquad
  \frg = f\pot\M\pot+f\dep\M\dep-\I,
\end{equation}
%
where $\I$ is the identity matrix.
Eventually, this will settle into the equilibrium distribution:
%
\begin{equation}\label{eq:eqprob}
  \eq\frg=0.
\end{equation}
%

@@

The \KO\ mice have a lower threshold for depression.
We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use a larger value for $q$ for depression in the \KO\ mice.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use a larger value for $x$ for depression in the \KO\ mice.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use larger values for depression in the \KO\ mice.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{tab:params}.

@@

This model is parameterised by the values of $x$ chosen for potentiation and depression.
We will use a larger value for depression in the \KO\ mice.
The values used are listed in \autoref{tab:params}.

@@

Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
It will change the environment, which will lead to a change in how often potentiation and depression occur.
It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
This approach was used to model motor learning in \cite{Smith2006savings}.

@@

We will assume $f\dep\wt=f\dep\ko$.
This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
Therefore the rates of potentiation and depression should not change very much either.
This implies that the mean synaptic weight in equilibrium differs between the \KO\ mice and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and \KO\ mice for the reasons discussed above.
We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{tab:params}.

We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.

@@

The features of the actual experiments that we'd like to capture are:
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the \KO\ mice.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the \KO\ mice, gain-increase learning is significantly faster with pre-training than without it.
  \item After pre-training, gain-increase learning is significantly faster in the \KO\ mice than in the wild-type.
  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
\end{enumerate}
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.

We will not worry too much about the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ mice or with/without pre-training, it can be ignored.

@@

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
    \cline{1-7}
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
%    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
%    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
%                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
    Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
    Non-uni. MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mice with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mice,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mice without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. }

@@

Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ mice, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.

However, this is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ mice.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
This happens sooner for the \KO\ mice than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ mice, even outside the desired range of values for $\Delta f$.
As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ mice have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ mice learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ mice but slows it down in the wild-type.
However, in this case pre-training proceeds much faster in the \KO\ mice than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ mice, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ mice than wild-type.

Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ mice (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
This will last for a shorter time in the \KO\ mice, due to the enhanced transitions.
This will not be a feature of any of the other models that we study here.
Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
The difference in these two curvatures is therefore a good feature of this model.
